### 2018-09-04 Manuscripts/cog-psych/revision/_region_.tex
Primary revisions to manuscript: copytweak-rep in section 8.5, unsup cat, contrast in thomas98, some mention of machine learning generative models. In other news, updated and more or less finalised fits with the category [0] bug fixed in Simulation.py. TODO: Update the goldstate.

%!TEX output_directory = latex_out/
\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage{csquotes}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage[toc]{appendix}
\usepackage[natbibapa]{apacite}

@@

@@

% ------- ABSTRACT PAGE ------- %
\doublespacing
\section*{Abstract}
\jlanote{need to update} 
\sxlnote{updated} 
%
The ability to creatively generate new concepts and ideas is among the most
fascinating aspects of human cognition, but we do not have a strong
understanding of the cognitive processes and representations underlying concept
generation. In this paper, we study the generation of new categories using the
computational and behavioral toolkit of traditional artificial category
learning. Previous work in this domain has focused on how the statistical
structure of known categories generalizes to generated categories, overlooking
the degree of contrast between the known and generated categories as a factor.
We report three experiments demonstrating that contrast between what is known
and what is created is of fundamental importance for categorization and the
creative process. We propose two novel approaches to modeling category contrast:
one focused on exemplar dissimilarity and another on the representativeness
heuristic. These new models are compared to two key alternatives. Our
experiments and simulations demonstrate specifically how category contrast
influences creative generation, and how it can interact with other constraints
to produce different types of creations. Our work also serves as a concrete
example of how well-established and highly controlled experimental frameworks
(such as those in the field of category learning) can be used to make progress
in an intriguing yet understudied domain.

@@

Previous work in category learning and generation has established that people
are highly sensitive to the structural properties of categories, such as
correlations between the features of category members and the relation between
items within the same category and those in different categories
\citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.
Inspired by this work, previous research on the topic of category generation has
explored a similar principle: People tend to create new categories that have
similar {\em statistical regularities} as previously learned categories
\citep{jern2013probabilistic,ward1994structured}. Although this is an important
characteristic of generating new categories, it cannot be the only one. Taken to
the extreme, the best ``new'' category in terms of having the same statistical
regularities to other categories would be identical to a known category that is
representative of the domain (and thus, not new at all).

\sxlnote{I think we can address Charles' point on real world examples of how
  contrast can be important around here.}
To successfully generate something novel, what is generated must be different
from what is already known. This fundamental constraint, ``being different'', or
contrasting from other categories in the relevant domain, is the focus of our
work. Although implicitly assumed in some work, this constraint has been
overlooked in previous research: To our knowledge, there has not been any
systematic investigation addressing how generated categories {\em differ} from
what is already known. Although the idea of category contrast is ubiquitous
throughout the categorization literature, and extends to a variety of other
fields \citep[e.g., color;][]{regier2007}, the idea that a new category should
be ``different'' is vague, as there are many ways it could be different from a
previously observed category.

@@

The second hypothesis for how contrast might affect categorization is contrast
as {\em representativeness}. The representativeness heuristic \citep{kahneman73}
states that people make judgments regarding an outcome based on how
representative it is of the evidence given in the current context. The heuristic
is a powerful theoretical construct that has been used to capture a range of
complex patterns of human judgments \citep{kahneman73,tversky74,tversky83},
especially those that deviate from normative theory (as given by a
straightforward application of Bayes' rule). For example, the coinflip sequence
TTHTH is perceived to be more random (due to it being representative of a
coinflip sequence) than TTTTT (despite both having the same probability of
having been generated from a fair coin). Although there is a healthy debate as
to whether perceived randomness {\em really} deviates from normative theory
\citep{griffiths18,hahn09}, the representativeness heuristic remains one of the
main explanations of human judgments \citep{reimers18}. To formalize the
hypothesis that people expect a new category to be representative of the
opposite of the current category, or contrast as {\em representativeness}, we
use the Bayesian formulation of representativeness \citep{tenenbaum01}, which
has been used successfully to capture human performance in color categorization
\citep{abbott16}, image category learning \citep{abbott11}, and language grammar
learning \citep{rafferty10}. We do so by extending the state-of-the-art category
generation model \citep{jern2013probabilistic} and find that human category
generation is best captured as generating representative, rather than probable,
samples.

@@

\sxlnote{Added thomas98 description here} It is worth pointing out that the
observation of regularities in distributional properties across categories is
not confined to the work of Ward and colleagues. Most notably, \citet{thomas98}
found that when people were asked to generate feature values for exemplars with
a missing feature (specifically circles of different sizes, each with a radial
line of varying orientation), the resulting set of exemplars tended to share the
same feature correlations as exemplars from previously learned categories. [the
transition to the next point isn't great, though]

@@

@@

\jlanote{check to make sure the next paragraph has been updated enough}
\sxlnote{I think it's updated enough! If given a chance I might edit this a
  little to be less harsh on hierarchical bayesian models in general though I
  think, since the representativeness model is one of them?}
Although it is evident that people are {\em capable} of creating new concepts
and categories, it is not entirely clear how new concepts are systematically
made different from what is already known. The hierarchical Bayesian model
developed by \cite{jern2013probabilistic} assumes that differences between
observed and generated categories are only due to random variation. The model
assumes that generated categories are sampled from the same underlying domain
distribution as observed categories, and will thus share a common distributional
structure. The model does not make predictions about the {\em location} of the
category within the domain (the perceptual instantiation of category members).
Under a strict interpretation of their model, given knowledge of a single
category within the domain, the most probable new category to be generated is
located in {\em exactly} the same location and possesses an identical
distributional structure. This is not an issue with their model specifically,
but with a broader class of standard hierarchical Bayesian models \citep[e.g.,
][]{griffithsscn07,kemp07}. Many of these models assume that at some point of
the latent generative process the same underlying distribution generates all of
the categories and thus, any differences between categories are due to {\em
noise} and should not be {\em systematic}. If the goal of creating a new
category given others in a domain is generating from the posterior predictive
distribution then the best a standard hierarchical Bayesian model without a
notion of contrast built into the model can do at capturing contrast is to
assume that the new category is placed uniformly at random over stimulus space.
But, this defeats the purpose of a hierarchy as it is ignored when determining a
new category location!\footnote{It is plausible that some hierarchical Bayesian
model could be created that generates categories different from each other.
However, this model would not be a standard application or extension of most
pre-existing hierarchical Bayesian models. The generative process would need to
include a component that promotes contrast.} Note that this does not mean
location information and contrast cannot be captured using hierarchical Bayesian
models. In fact, our model of contrast as representativeness is a hierarchical
Bayesian model, but with a different goal than predicting the next category.

@@

\sxlnote{Charles suggested we either remove the section below, or move it to the
  discussion. I've tried finding a good place for it in the discussion but it
  doesn't fit very well, and moving it takes away this nice motivation for
  our further investigation into contrast. I think that keeping it here in the
  introduction might work out, provided Charles finds my argument for why the
  regularity assumption is too strong convincing enough. More details a little
  further down.}
\subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}

\sxlnote{Added this bit on regularity below - I'm a little undecided as to leave
  it in the main text or place as a gigantic footnote? It is a bit of a
  digression, and it's really just here to address Charles' point.... If we want
  to keep this in the main text, I'll have to amend the following paragraph
  ``Re-analyzing the results...'' to improve the transition.}
%
It is possible, however, that this data can be explained by a process that does
not involve contrast. Specifically, since participants were trained on exemplars
that share the same hue within a category, it is plausible that they generate
exemplars that all share a common hue. If this is the case, given a $6/8 = .75$
probability that a hue distinct from the experimenter-defined categories is
selected, there is a corresponding $.75$ probability (or about 17 out of 22
participants) that the generated category will have no exemplars that share a
hue with any of the experimenter-defined categories. In addition, that there is
a $2/8 = .25$ probability (about 6 out of 22 participants) that the generated
category will have all six exemplars that share a hue with any of the
experimenter-defined categories. This prediction of regularity across categories
is much closer to the data than the prediction from a Binomial model and does
not require any mechanism of category contrast.

@@

Despite both the original hierarchical Bayesian model and our representativeness
model sharing identical hierarchical structures, their distinct response
processes can yield very different exemplars. Specifically, the hierarchical
Bayesian model emphasizes the generation of exemplars that maintain
distributional commonalities across categories. In contrast, the
representativeness model focuses on the generation of exemplars that are
representative of the underlying distribution for a target category. Generally,
this mechanism of representativeness results in the generation of exemplars that
are less similar to exemplars from other categories. We see this illustrated in
Figure \ref{fig:representative-examples}, where the representativeness model
displays a strong preference for generating exemplars from target category 'B'
that are further away from the contrast category 'A'.

@@

\section{Model-based Analyses}
\sxlnote{For this section I've added the existence of the representativeness
  model but I think discussion of its performance is still lacking compared to PACKER.}
Experiments 1 and 2 revealed systematic and strong effects of category contrast
on creative generation. In this section, we report the results of simulations
with formal models aimed at explaining our observations. Specifically, we
present simulations from the two novel contrast models: PACKER and the
representativeness model, and compare them to two models that do not incorporate
contrast: the copy-and-tweak model (discussed in Section
\ref{section:copytweak-packer}) and an implementation of the hierarchical
Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
Appendix \ref{ap:hsampling-definition}. For our simulations here, the
copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
parameter constrained to be zero. The comparison of this set of models serves to
highlight the explanatory role of contrast in categorization: If contrast
affords little explanatory advantage, then there should be few differences in
performance between PACKER and copy-and-tweak, or between the representativeness
and hierarchical Bayesian model. The comparison between these models can also
emphasize the necessity of contrast and demonstrate that generation cannot be
explained entirely through the emulation of distributional structure. Each model
has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
are insensitive to the distributional structure of learned categories (relying
only on exemplar similarities), the representativeness and hierarchical Bayesian
model generates categories exclusively on the basis of knowledge of how existing
classes are distributed.

@@

\sxlnote{The argument in the following two paragraphs appear to be challenged by
  the success of the representativeness model. I might suggest removing them altogether?}
While PACKER's advantage may tentatively be attributed to the model's
sensitivity to category contrast (this will be explored in detail below), the
advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
attributed to its exemplar-based representation, as opposed to the
prototype-based representation assumed by the hierarchical Bayesian model. As
observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
categories we observed were often widely distributed, with no items near the
category prototype. This aspect of the data is inconsistent with the
multivariate normal distributions (similar to prototypes) used to represent
categories in the \cite{jern2013probabilistic} model, but can be easily
accounted for using an exemplar-based approach.

A key distinction between PACKER and the other three models is that only PACKER
is capable of making strong predictions about the location of new category
members when the target class is entirely novel (i.e., no member of the category
has been observed). Under these circumstances, there are no examples to copy,
and thus the copy-and-tweak model predicts that items are generated at random.
Likewise, with no observations on which to condition the category distribution,
the hierarchical Bayesian and representativeness models also picks an item at
random. Crucially, however, the next generated exemplar from the
representativeness model is strongly influenced by contrast as
representativeness. Thus, it is possible that the failure of these models is
simply due to their inability to explain each participant's first trial
(generating the first item in the `Beta' category). To ensure this is not
driving our results, we conducted an identical set of simulations as above,
excluding this trial (leaving 915 responses in the dataset): Again, PACKER
($L = -3390$, $AIC = 6786$) achieved better fits than the copy-and-tweak
($L = -3579$, $AIC = 7162$) and hierarchical Bayesian ($L=-3612$, $AIC = 7232$)
models.

Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
ratio test to compare the two models. PACKER explains the aggregate data
significantly better than copy-and-tweak ($\chi^2(1) = 749, p < 0.001$ for all
data and $\chi^2(1) = 377, p < 0.001$ excluding the first example), providing
further evidence that category generation is better explained when contrast is
considered.

@@

As in PACKER, changes in the parameter settings of the copy-and-tweak model can
also be used to produce different patterns of generation. Indeed, as
copy-and-tweak is simply a special case of the PACKER model, the attention
weights operate exactly as described above to produce row- and column-like
categories. However, because the model is not influenced by category contrast,
it is biased toward generating tightly clustered categories, as new items are
always most likely to be generated near known examples of the target category.
Thus, the lack of a contrast mechanism prevents the model from explaining why
some individuals widely distribute their categories to the corners of the space.

Similar to the copy-and-tweak model, the hierarchical Bayesian model possesses
no mechanism to account for category contrast, and the model is most likely to
generate new items that are similar to known examples of the target category.
Although this precludes an account of why individuals might assume a cluster
versus corners profile, the covariance matrix specifying the model's prior
domain distribution, $\Sigma_0$, can be used to explain the generation of
row-like and column-like categories. This covariance matrix specifies the amount
of variance assumed along each dimension (as well as the dimensional
correlations) across the domain of categories. The covariance matrix for a newly
generated category, $\Sigma_B$, is based on the assumed $\Sigma_0$ as well as
the distributions of previously learned categories (see Appendix
\ref{ap:hsampling-definition}). Thus, the importance of each feature can be
coded into $\Sigma_0$ to alter the dimensional variance of generated categories.

@@

\section{General Discussion}

The creative generation of concepts is an immensely intriguing topic, but it is
also complex and difficult to study in a traditional laboratory setting.
Creative generation is most typically studied using a creative cognition
approach \citep{finke1992creative,smith1995creative}, where creative products
\citep[such as drawings of alien plants and animals;][]{ward1994structured} are
analyzed to obtain insight into the processes and representations involved.
While this approach has yielded a great deal of information about the nature of
creative generation, the commonly employed response modalities (e.g., hand-drawn
pictures) make it difficult to develop computational models capable of formally
evaluating the proposed representations and processes. However, the generation
of novel concepts is not altogether different from other behaviors studied in
cognitive psychology: In particular, it can be viewed as an form of inference
using one's knowledge about existing categories
\citep{kemp2014taxonomy,kurtz2015human}. Taking advantage of this insight, in
this paper we investigated the generation of new categories using the empirical
and computational toolkit from the field of categorization.

While the bulk of prior research on the topic has focused on the classic finding
that generated concepts tend to be distributionally similar to known concepts,
there has been little work addressing the role of contrast in creative
generation: How is it that people are able to create something {\em different}
from what is already known? We developed a novel, exemplar-based model, PACKER,
which formally specifies the role of contrast in generation. The model proposes
that categories are represented as exemplars in a multidimensional psychological
space, and generation is constrained both by within-category and
between-category similarity: Exemplars belonging to the same category should be
similar to one another, and exemplars belonging to different categories should
not be similar to one another.

In addition to an analysis of published data \citep[][Experiment
3]{jern2013probabilistic}, we reported two experiments demonstrating systematic
effects of category contrast in creative generation. Members of
participant-generated categories tended to be highly dissimilar from members of
previously-learned categories, and were usually more similar to one another than
to members of other categories. We also observed broad interdependence between
the distributional structure (feature variance, correlation) and physical
instantiation (location within the stimulus space) of generated categories: In
Experiment 2, we found that the unoccupied regions of the domain influenced the
distributional structure of categories, and in both experiments we observed that
participants distributed their generated categories to increase contrast with
what was already known.

We conducted simulations comparing PACKER's account of our results to the other
major proposals for category generation: a ``copy-and-tweak'' model (realized as
a variant of PACKER with no sensitivity to category contrast), and a
hierarchical Bayesian model designed to explain the classic distributional
similarity effect. In all simulations, we found that PACKER's sensitivity to
contrast captured a previously unexplained and unexplored aspect of human
category generation. Further, by measuring PACKER's fit as a function of its
prioritization of within- and between-category similarity, we observed that
considering either constraint exclusively results in a relatively low-quality
account. Instead, PACKER's best results were obtained when both constraints are
considered, indicating that human learners do not generate novel concepts
exclusively on the basis of within-category similarity or between
class-contrast. This finding mirrors our behavioral results and demonstrates
that both constraints influence creative generation.

@@

\subsection{Implications for Creative Cognition}

Although the focus of this article has been to address the processes involved in
creative generation using the empirical and quantitative toolkit of traditional
categorization research, our findings and approach have relevant implications
for research in creative cognition. A central focus of the creative cognition
approach has been to explain acts of creativity in terms of the mental
representations and processes that are commonly studied in cognitive psychology
and cognitive science \citep{finke1992creative,smith1995creative}. However,
unlike other fields in the study of cognition, creative cognition research
rarely employs quantitative models to evaluate the explanatory value of such
representations and processes. Our modeling results provide a concrete example
of how formal approaches may be used to gain insight into the nature of creative
cognition.

>>>>

\message{ !name(cogpsych.tex)}%!TEX output_directory = latex_out/
\documentclass[12pt]{article}
\usepackage[letterpaper, margin=1in, headheight=15pt]{geometry}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pgfplots}
\usepackage{fancyhdr}
\usepackage{csquotes}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage[toc]{appendix}
\usepackage[natbibapa]{apacite}

@@

\message{ !name(cogpsych.tex) !offset(-3) }

@@

% ------- ABSTRACT PAGE ------- %
\doublespacing
\section*{Abstract}
%
The ability to creatively generate new concepts and ideas is among the most
fascinating aspects of human cognition, but we do not have a strong
understanding of the cognitive processes and representations underlying concept
generation. In this paper, we study the generation of new categories using the
computational and behavioral toolkit of traditional artificial category
learning. Previous work in this domain has focused on how the statistical
structure of known categories generalizes to generated categories, overlooking
the degree of contrast between the known and generated categories as a factor.
We report three experiments demonstrating that contrast between what is known
and what is created is of fundamental importance for categorization and the
creative process. We propose two novel approaches to modeling category contrast:
one focused on exemplar dissimilarity and another on the representativeness
heuristic. These new models are compared to two key alternatives. Our
experiments and simulations demonstrate specifically how category contrast
influences creative generation, and how it can interact with other constraints
to produce different types of creations. Our work also serves as a concrete
example of how well-established and highly controlled experimental frameworks
(such as those in the field of category learning) can be used to make progress
in an intriguing yet understudied domain.

@@

Previous work in category learning and generation has established that people
are highly sensitive to the structural properties of categories, such as
correlations between the features of category members and the relation between
items within the same category and those in different categories
\citep{roschmervis1975,regier2007,shepard1961learning,smith1995creative}.
Inspired by this work, previous research on the topic of category generation has
explored a similar principle: People tend to create new categories that have
similar {\em statistical regularities} as previously learned categories
\citep{jern2013probabilistic,ward1994structured}. Although this is an important
characteristic of generating new categories, it cannot be the only one. Taken to
the extreme, the best ``new'' category in terms of having the same statistical
regularities to other categories would be identical to a known category that is
representative of the domain (and thus, not new at all). Further, contrast has
played a critical role studying the spread of new ideas in diverse fields, such
as marketing \citep{berger16} and sociology \citep{rogers03}.

To successfully generate something novel, what is generated must be different
from what is already known. This fundamental constraint, ``being different'', or
contrasting from other categories in the relevant domain, is the focus of our
work. Although implicitly assumed in some work, this constraint has been
overlooked in previous research: To our knowledge, there has not been any
systematic investigation addressing how generated categories {\em differ} from
what is already known. Although the idea of category contrast is ubiquitous
throughout the categorization literature, and extends to a variety of other
fields \citep[e.g., color;][]{regier2007}, the idea that a new category should
be ``different'' is vague, as there are many ways it could be different from a
previously observed category.

@@

The second hypothesis for how contrast might affect categorization is contrast
as {\em representativeness}. The representativeness heuristic \citep{kahneman1972subjective}
states that people make judgments regarding an outcome based on how
representative it is of the evidence given in the current context. The heuristic
is a powerful theoretical construct that has been used to capture a range of
complex patterns of human judgments \citep{kahneman73,tversky74,tversky83},
especially those that deviate from normative theory (as given by a
straightforward application of Bayes' rule). For example, the coinflip sequence
TTHTH is perceived to be more random (due to it being representative of a
coinflip sequence) than TTTTT (despite both having the same probability of
having been generated from a fair coin). Although there is a healthy debate as
to whether perceived randomness {\em really} deviates from normative theory
\citep{griffiths18,hahn09}, the representativeness heuristic remains one of the
main explanations of human judgments \citep{reimers18}. To formalize the
hypothesis that people expect a new category to be representative of the
opposite of the current category, or contrast as {\em representativeness}, we
use the Bayesian formulation of representativeness \citep{tenenbaum01}, which
has been used successfully to capture human performance in color categorization
\citep{abbott16}, image category learning \citep{abbott11}, and language grammar
learning \citep{rafferty10}. We do so by extending the state-of-the-art category
generation model \citep{jern2013probabilistic} and find that human category
generation is best captured as generating representative, rather than probable,
samples.

@@

@@

It is worth pointing out that the early observation of regularities in
distributional properties across categories is not confined to the work of Ward
and colleagues. Most notably, \citet{thomas98} trained participants to classify
circles of different sizes, each with a radial line of varying orientation, into
two categories. During a surprise prediction phase, participants were asked to
generate a value for a missing feature for a certain category exemplar given a
value on the other feature. Their results revealed that exemplars tended to
share the same feature correlations as exemplars from previously learned
categories. Interestingly, this effect was not consistently observed when the
given values of a feature fell outside the range of learned values for that
particular category. For instance, if the learned Category A exemplars were
generally 8 cm to 10 cm in radius and negatively correlated with the angle of
the radial line, participants generally produced radial line orientations that
were also negatively correlated with the exemplar size when the given size was
between 8 cm and 10 cm. However, when the given size fell outside that range for
that category (e.g., 14 cm for a Category A exemplar), some participants
consistently produced radial line orientations that were not negatively
correlated with the size and that also placed the exemplars in a significantly
different location in the feature space. These results suggest that when tasked
to produce exemplars that fall outside what was previously learned, people show
some tendency to generate exemplars that are different (both distributionally
and spatially) to what was learned. We explore this idea of contrast
more deeply in the following section.

@@

Although it is evident that people are {\em capable} of creating new concepts
and categories, it is not entirely clear how new concepts are systematically
made different from what is already known. The hierarchical Bayesian model
developed by \cite{jern2013probabilistic} assumes that differences between
observed and generated categories are only due to random variation. The model
assumes that generated categories are sampled from the same underlying domain
distribution as observed categories, and will thus share a common distributional
structure. The model does not make predictions about the {\em location} of the
category within the domain (the perceptual instantiation of category members).
Under a strict interpretation of their model, given knowledge of a single
category within the domain, the most probable new category to be generated is
located in {\em exactly} the same location and possesses an identical
distributional structure. This is not an issue with their model specifically,
but with a broader class of standard hierarchical Bayesian models \citep[e.g.,
][]{griffithsscn07,kemp07}. Many of these models assume that at some point of
the latent generative process the same underlying distribution generates all of
the categories and thus, any differences between categories are due to {\em
noise} and should not be {\em systematic}. If the goal of creating a new
category given others in a domain is generating from the posterior predictive
distribution then the best a standard hierarchical Bayesian model without a
notion of contrast built into the model can do at capturing contrast is to
assume that the new category is placed uniformly at random over stimulus space.
But, this defeats the purpose of a hierarchy as it is ignored when determining a
new category location!\footnote{It is plausible that some hierarchical Bayesian
model could be created that generates categories different from each other.
However, this model would not be a standard application or extension of most
pre-existing hierarchical Bayesian models. The generative process would need to
include a component that promotes contrast.} Note that this does not mean
location information and contrast cannot be captured using hierarchical Bayesian
models. In fact, a model instantiating contrast as representativeness is a
hierarchical BAyesian model. But, it requires an additional factor, which in our
case is having a different goal than predicting the next category.

@@

@@

Despite both the original hierarchical Bayesian model and our representativeness
model sharing identical hierarchical structures, their distinct response
processes can yield very different exemplars. Specifically, the hierarchical
Bayesian model emphasizes the generation of exemplars that maintain
distributional commonalities across categories. In contrast, the
representativeness model focuses on the generation of exemplars that are
representative of the underlying distribution for a target category.
\jlanote{The next sentence is weird sounding and should be cleaned up.}
\sxlnote{Ok! I hope it reads better now.} Generally, this mechanism of
representativeness results in the generation of exemplars that are less similar
to exemplars from other categories. We see this illustrated in Figure
\ref{fig:representative-examples}, where the representativeness model displays a
strong preference for generating exemplars from target category 'B' that are
further away from the contrast category 'A'.

@@

\section{Model-based Analyses}
Experiments 1 and 2 revealed systematic and strong effects of category contrast
on creative generation. In this section, we report the results of simulations
with formal models aimed at explaining our observations. Specifically, we
present simulations from the two novel contrast models: PACKER and the
representativeness model, and compare them to two models that do not incorporate
contrast: the copy-and-tweak model (discussed in Section
\ref{section:copytweak-packer}) and an implementation of the hierarchical
Bayesian model proposed by \cite{jern2013probabilistic}, described in-depth in
Appendix \ref{ap:hsampling-definition}. For our simulations here, the
copy-and-tweak model is defined as a variant of PACKER with the $\theta_c$
parameter constrained to be zero. The comparison of this set of models serves to
highlight the explanatory role of contrast in categorization: If contrast
affords little explanatory advantage, then there should be few differences in
performance between PACKER and copy-and-tweak, or between the representativeness
and hierarchical Bayesian model. The comparison between these models can also
emphasize the necessity of contrast and demonstrate that generation cannot be
explained entirely through the emulation of distributional structure. Each model
has complementary strengths and weaknesses: Whereas PACKER and copy-and-tweak
are insensitive to the distributional structure of learned categories (relying
only on exemplar similarities), the representativeness and hierarchical Bayesian
model generates categories exclusively on the basis of knowledge of how existing
classes are distributed.

@@

While PACKER's advantage may tentatively be attributed to the model's
sensitivity to category contrast (this will be explored in detail below), the
advantage shown by copy-and-tweak over the hierarchical Bayesian model may be
attributed to its exemplar-based representation, as opposed to the
prototype-based representation assumed by the hierarchical Bayesian model. As
observed in Figures \ref{fig:e1-samples} and \ref{fig:e2-samples}, the generated
categories we observed were often widely distributed, with no items near the
category prototype. This aspect of the data is inconsistent with the
multivariate normal distributions (similar to prototypes) used to represent
categories in the \cite{jern2013probabilistic} model, but can be easily
accounted for using an exemplar-based approach.

A key distinction between the contrast and classical models is that only the
contrast models are capable of making strong predictions about the location of
new category members when the target class is entirely novel (i.e., no member of
the category has been observed). Under these circumstances, there are no
examples to copy, and thus the copy-and-tweak model predicts that items are
generated at random. Likewise, with no observations on which to condition the
category distribution, the hierarchical Bayesian model also picks an item at
random.

Thus, it is possible that the failure of the classical models is simply due to
their inability to explain each participant's first trial (generating the first
item in the `Beta' category). To ensure this is not driving our results, we
conducted an identical set of simulations as above, excluding this trial
(leaving 915 responses in the dataset): Again, the representativeness model ($L
= -3286$, $AIC = 6580$) and PACKER ($L = -3377$,
$AIC = 6759$) achieved better fits than the copy-and-tweak ($L = -3564$,
$AIC = 7132$) and hierarchical Bayesian ($L = -3597$, $AIC = 7201$) models.

Finally, because copy-and-tweak is nested within PACKER, we can use a likelihood
ratio test to compare the two models. PACKER explains the aggregate data
significantly better than copy-and-tweak ($\chi^2(1) = 747, p < 0.001$ for all
data and $\chi^2(1) = 375, p < 0.001$ excluding the first example), providing
further evidence that category generation is better explained when contrast is
considered.

@@

As in PACKER, changes in the parameter settings of the copy-and-tweak model can
also be used to produce different patterns of generation. Indeed, as
copy-and-tweak is simply a special case of the PACKER model, the attention
weights operate exactly as described above to produce row- and column-like
categories. However, because the model is not influenced by category contrast,
it is biased toward generating tightly clustered categories, as new items are
always most likely to be generated near known examples of the target category.
Thus, the lack of a contrast mechanism prevents the model from explaining why
some individuals widely distribute their categories to the corners of the space.
\jlanote{Xian: Please add a paragraph or two on the representativeness model
  here.}
\sxlnote{Ok! I've kind of integrated it into one bigger paragraph with
  hierarchical bayes.}

Unlike PACKER, both the hierarchical Bayesian model and the representative model
do not contain any explicit mechanism for controlling the weights of
between-category contrast and within-category similarity. This means both models
do not make strong predictions of cluster versus corner profiles. However, the
covariance matrix specifying the model's prior domain distribution, $\Sigma_0$,
can be used to explain the generation of row-like and column-like categories.
This covariance matrix specifies the amount of variance assumed along each
dimension (as well as the dimensional correlations) across the domain of
categories. The covariance matrix for a newly generated category, $\Sigma_B$, is
based on the assumed $\Sigma_0$ as well as the distributions of previously
learned categories (see Appendix \ref{ap:hsampling-definition}). Thus, the
importance of each feature can be coded into $\Sigma_0$ to alter the dimensional
variance of generated categories. Because the hierarchical Bayesian model
possesses no mechanism to account for category contrast, the model is most
likely to generate new items that are similar to known examples of the target
category with no regard for how different it is to the contrast category.
However, the representativeness model predicts that new exemplars should be more
representative (i.e., more similar to target and less similar to contrast
categories), accounting for the tendency of row-like and column-like profiles
occupying the edge of the feature space.

@@

\section{General Discussion}
The sensory impression of every stimulus and event is unique. Grouping distinct
patterns of sensory information into categories is a fundamental task solved by
the mind. Most work has focused on how people learn new categories that are
provided to them through unlabeled, partially labeled, or fully-labeled
examples. How were these categories first determined? Some natural categories
are likely to be the result of regularities in the dynamics of our environment.
But, these are only a subset of the categories that people learn. Other
categories, such as tools and ideas, were generated by people over time. What
basic principles underlie how people generate categories?

% The creative generation of concepts is an immensely intriguing topic, but it is
% also complex and difficult to study in a traditional laboratory setting.
% Creative generation is most typically studied using a creative cognition
% approach \citep{finke1992creative,smith1995creative}, where creative products
% \citep[such as drawings of alien plants and animals;][]{ward1994structured} are
% analyzed to obtain insight into the processes and representations involved.
% While this approach has yielded a great deal of information about the nature of
% creative generation, the commonly employed response modalities (e.g., hand-drawn
% pictures) make it difficult to develop computational models capable of formally
% evaluating the proposed representations and processes. However, the generation
% of novel concepts is not altogether different from other behaviors studied in
% cognitive psychology: In particular, it can be viewed as an form of inference
% using one's knowledge about existing categories
% \citep{kemp2014taxonomy,kurtz2015human}. Taking advantage of this insight, in
% this paper we investigated the generation of new categories using the empirical
% and computational toolkit from the field of categorization.

While the bulk of prior research on the topic has focused on the classic finding
that generated concepts tend to be distributionally similar to known concepts,
there has been little work addressing the role of contrast in creative
generation: How is it that people are able to create something {\em different}
from what is already known?
\jlanote{Update this next section:}
\sxlnote{Updated}
 We developed two novel models, each incorporating a
different conceptualization of category contrast. One model is PACKER, which is
an exemplar-based model, that formally specifies the role of contrast in
generation as exemplar dissimilarity. Specifically, the model proposes that
categories are represented as exemplars in a multidimensional psychological
space, and generation is constrained both by within-category and
between-category similarity: Exemplars belonging to the same category should be
similar to one another, and exemplars belonging to different categories should
not be similar to one another. The second model is a novel hierarchical
Bayesian model involving a representativeness mechanism. This model generates
exemplars that are more representative of (i.e., has more evidence from) the
novel category compared to the learned category.

We reported two experiments demonstrating systematic effects of category
contrast in creative generation. Members of participant-generated categories
tended to be highly dissimilar from members of previously-learned categories,
and were usually more similar to one another than to members of other
categories. We also observed broad interdependence between the distributional
structure (feature variance, correlation) and physical instantiation (location
within the stimulus space) of generated categories: In Experiment 2, we found
that the unoccupied regions of the domain influenced the distributional
structure of categories, and in both experiments we observed that participants
distributed their generated categories to increase contrast with what was
already known.

We conducted simulations comparing the contrast models' account of our results
to the classical proposals for category generation: a ``copy-and-tweak'' model
(realized as a variant of PACKER with no sensitivity to category contrast), and
a hierarchical Bayesian model designed to explain the classic distributional
similarity effect. In all simulations, we found that the contrast models captured a previously unexplained and unexplored aspect of human
category generation. In particular, by measuring PACKER's fit as a function of its
prioritization of within- and between-category similarity, we observed that
considering either constraint exclusively results in a relatively low-quality
account. Instead, PACKER's best results were obtained when both constraints are
considered, indicating that human learners do not generate novel concepts
exclusively on the basis of within-category similarity or between
class-contrast. This finding mirrors our behavioral results and demonstrates
that both constraints influence creative generation when explained with an
exemplar model. Ultimately, our category generation data was best explained by
the representativeness model. The exact nature of its superiority over the other
models is still not yet clear. However, as we explore in a later section, it is
likely that the representativeness mechanism itself (i.e., independent of the
hierarchical Bayesian foundation) captures a core aspect of both category
generation and category learning.
\jlanote{Above needs some updating still. Also, do we want an outline of general
discussion paragraph?}
\sxlnote{Updated. I tried writing an outline, but found that points got a little
  repetitive after a bit, so I ditched that idea... Might have just been the way
I wrote it though.}

\subsection{Novel Analyses Demonstrating Contrast Effects in Prior Work}

It is possible, however, that this data can be explained by a process that does
not involve contrast. Specifically, since participants were trained on exemplars
that share the same hue within a category, it is plausible that they generate
exemplars that all share a common hue. If this is the case, given a $6/8 = .75$
probability that a hue distinct from the experimenter-defined categories is
selected, there is a corresponding $.75$ probability (or about 17 out of 22
participants) that the generated category will have no exemplars that share a
hue with any of the experimenter-defined categories. In addition, that there is
a $2/8 = .25$ probability (about 6 out of 22 participants) that the generated
category will have all six exemplars that share a hue with any of the
experimenter-defined categories. This prediction of regularity across categories
is much closer to the data than the prediction from a Binomial model and does
not require any mechanism of category contrast.

\jlanote{Xian: Please do a Chi-Squared test between what is expected based on
  the logic above and the observed frequencies mentioned below}
\sxlnote{I didn't do a chi-sq test because what appears below is not the
  observed frequencies that correspond to the logic above, but is instead just
  an indication of whether or not it is reasonable to assume the logic above
  (and the argument is that it is not reasonable.)}

@@

\subsection{Implications for Creative Cognition}
\jlanote{this needs updating for representativeness}
\sxlnote{Updated.}
Although the focus of this article has been to address the processes involved in
creative generation using the empirical and quantitative toolkit of traditional
categorization research, our findings and approach have relevant implications
for research in creative cognition. A central focus of the creative cognition
approach has been to explain acts of creativity in terms of the mental
representations and processes that are commonly studied in cognitive psychology
and cognitive science \citep{finke1992creative,smith1995creative}. However,
unlike other fields in the study of cognition, creative cognition research
rarely employs quantitative models to evaluate the explanatory value of such
representations and processes. Our modeling results provide a concrete example
of how formal approaches may be used to gain insight into the nature of creative
cognition.
